A data-driven model of nucleosynthesis with chemical tagging in a lower-dimensional latent space

### 2019-09-09 article/mcfa.tex
things
foo

@@

@@

\begin{abstract}
Chemical tagging seeks to identify unique star formation sites from
present-day stellar abundances.
Previous techniques have treated each abundance dimension as being
statistically independent, despite theoretical expectations that
many elements can be produced by more than one nucleosynthetic process.
In this work we introduce a data-driven model of nucleosynthesis 
where a set of latent factors (e.g., nucleosynthetic yields) contribute
to all stars with different scores, and clustering (e.g., chemical tagging) 
is modelled by a mixture of multivariate Gaussians
in a lower-dimensional latent space.
We use an exact method to simultaneously estimate the factor scores for
each star, the partial assignment of each star
to each cluster, and the latent factors common to all stars, even in the
presence of missing data entries.
We use an information-theoretic Bayesian principle to estimate the number of
latent factors and clusters.
Using the second \Galah\ data release we find that five latent factors are
preferred to explain $N =$ 1,072 stars with 18 chemical abundances. 
We identify the rapid- and slow-neutron capture 
processes, as well as latent factors consistent with Fe-peak and 
$\alpha$-element production, and another where only K and Zn dominate.
When we consider $N \sim$ 100,000 stars with missing abundances we find 
another seven factors, as well as 16 components in latent space.
Despite these components showing separation in chemistry that is explained
through different yield contributions, none show significant structure in 
their positions or motions. We argue that more data, and joint priors 
on cluster membership that are constrained by dynamical models, are necessary
to realise chemical tagging at a galactic-scale.
We release accompanying software that scales well with the available data,
allowing for model parameters to be optimised in seconds given a fixed number
of latent factors, components, and $\sim10^7$ abundance measurements.
\end{abstract}

@@

Gravitationally bound star clusters have been useful laboratories for
testing the limits and utility of chemical tagging. Although biases arise when
only considering star clusters that are still gravitationally bound, the chemical
homogeneity of open clusters provides an empirical measure of how similar stars
would need to be before they could be tagged as belonging to the same
star formation site \citep{Bland-Hawthorn:2010b,Bland-Hawthorn:2010a,Mitschang:2014}. However, there are analysis
issues in understanding how precisely those chemical abundances can be measured
\citep{Bovy:2016}, and how chemically similar stars can be that did not form 
together \citep[doppleg\"angers;][]{Ness:2018}.
If open clusters were truely chemically homogeneous then under idealistic 
assumptions our ability to chemically tag the Milky Way would depend primarily
on the precision with which we can measure those chemical abundances in stars. 
Data-driven approaches to modelling stellar spectra are
improving upon this precision \citep{Ness:2015,Ness:2018a,Ness:2018b,
Casey:2016,Casey:2017,Ho:2017b,Ho:2017a,Leung;Bovy:2018}, but more work is
needed: astronomers have not yet developed unbiased estimators of chemical
abundances that saturate the Cram\'er-Rao bound \citep{Cramer:1946,Rao:1945}.

@@

Given these data and the most favourable assumptions in chemical tagging 
-- that star clusters are truely chemically homogenous, that we can measure 
those abundances with infinite precision, and that those abundances are 
differentiable between star clusters -- then chemical
tagging becomes a clustering problem. All clustering techniques applied to 
chemical tagging thus far have assumed that the data dimensions are independent. That is to say
that adding a dimension of say [Ni/H] provides independent information
that could not have been predicted from other elemental abundances.
Theory and observations agree that this cannot be true.
Nucleosynthetic processes produce multiple elements in varying
quantities, and the effective dimensionality of stellar abundance datasets has been shown
to be lower than the actual number of abundance dimensions \citep{Ting:2012,Price-Jones:2018}.
Any clustering approach that treats each new elemental abundance as an 
independent axis of information will therefore conclude with biased inferences
about the star formation history of our Galaxy.

@@

Latent factor analysis is a common statistical approach for describing correlated 
observations with a lower number of latent variables \citep[e.g.,][]{Thompson:2004}.
Related techniques include principal component analysis \citep{Hotelling:1933} and its
variants \citep{Tipping;Bishop:1999}, singular value decomposition \citep{Golub:1970}, and other
matrix factorization methods. While factor analysis on its own is a useful
dimensionality reduction tool to identify latent factors that contribute to
the chemical abundances of stars \citep[e.g.,][]{Ting:2012,Price-Jones:2018}, factor
analysis cannot describe clustering in the data (or latent) space. As a result,
some works have performed clustering and then required different latent factors for each (totally assigned) component \citep[e.g.,][]{EdwardsDowe1998}.
Similarly, clustering techniques applied to chemical abundances to date 
\citep[e.g.,][]{Hogg:2016} do not account for the lower effective dimensionality in
elemental abundances.

@@

Here we introduce a toy model where we use generated data to verify that
we recover the true model parameters given some data, and to
ensure that the expectation-maximization method is yielding consistent results.
We generated a data set with ${\NumData = 1}$00,000 data points, each with
$\NumDimensions = 15$ dimensions. We adopted a latent dimensional space of 
$\NumLatentFactors = 5$ factor loads such that the vector $\factorloads$ has
shape $\NumDimensions \times \NumLatentFactors$,
with $\NumComponents = 20$ clusters in the latent space. 
We generated the random factor loads in the same way that we initialise the optimisation (Section~\ref{sec:initialisation}). The relative weights $\vec\weight$
are drawn from a multinomial distribution and the means of the clusters
in factor scores $\scoremeans$ are drawn from a standard normal
distribution. The off-diagonal entries in the covariance matrices in factor scores $\scorecovs$ are drawn from a gamma distribution $\scorecovs_{\numcomponents,i,i} \sim \vec\Gamma\left(1\right)$. The variance in 
each dimension $\specificvariance$ are also drawn $\specificvariance \sim \vec\Gamma\left(1\right)$.
The $\numdata$th data point (which belongs to the $\numcomponents$th cluster) is then
generated by drawing $\factorscores_{\numdata} \sim \mathcal{N}(\scoremeans_\numcomponents,\scorecovs_\numcomponents)$, projecting by the factor loads $\factorloads$, and adding variance $\specificvariance$.

@@

We treat the generated data set as if the number of latent factors
and components are not known. Starting with $\NumLatentFactors = 1$
and $\NumComponents = 1$, we trialled each permutation of $\NumLatentFactors$ and $\NumComponents$
until $\NumLatentFactors_\textrm{max} = 10$
and   $\NumComponents_\textrm{max} = 40$ (e.g., twice the true values of $\NumLatentFactors_\textrm{true}$ and $\NumComponents_\textrm{true}$).

We recorded the \emph{negative} log likelihood, the BIC, and the message length\footnote{Omitting constant terms such that negative message lengths are allowed.}  for each permutation of $\NumLatentFactors$ and $\NumComponents$.
These metrics are shown in Figure~\ref{fig:experiment-1-gridsearch}.
Unsurprisingly the negative log likelihood decreases with increasing numbers of latent
factors $\NumLatentFactors$ and increasing numbers of components $\NumComponents$.
The lowest BIC value and message length is found at $\NumLatentFactors = 5$
and $\NumComponents = 20$, identical to the true values.
It is clear from Figure~\ref{fig:experiment-1-gridsearch} that a combination of latent factors
and clustering in the latent space provides a better description of the (generated) data than a Gaussian mixture model without latent factors.
Adding components to the model does improves the log likelihood, even with a single latent factor,
but the addition of just \emph{one latent factor} improves the log likelihood more so than adding
\emph{twenty components}. Not much more can be said for this example because the true data generating process 
is known, but this toy model does illustrate how 
clustering in high dimensional data can be better described by latent factors with 
clustering in the lower dimensional latent space.

@@

We took the model with the preferred number of latent factors and components found
from a grid search ($\NumComponents = 20$, $\NumLatentFactors = 5$; which are also
the true values) and applied an orthogonal rotation to the latent space to be as
close as possible to the true values. The rotation matrix $\mathbf{R}$ was found
by solving for $\NumLatentFactors$ unknown angle parameters, each of which is used
to construct a Givens rotation matrix \citep{Givens:1958}, and then we take the product of those Givens
matrices to produce a valid rotation matrix $\vec{R}$. This process reduces to Euler angle rotation in three or fewer dimensions.
This process rotates the latent space
($\factorloads$, $\scoremeans$, $\scorecovs$), but has no effect on the model's 
predictive power: the evaluated log likelihood or the Kullback-Leibler divergence \citep{Kullback:1951} under the
rotated model is indistinguishable from the unrotated model.
In Figure~\ref{fig:exp1-compare} we show the estimated factor loads $\factorloads$,
factor scores $\factorscores$, and specific variances $\specificvariance$ compared
to the true values. The agreement is excellent in all model parameters.

@@

\begin{figure*}[t!]
	\includegraphics[width=\textwidth]{experiments/exp-missing-data-not-at-random-inflated-40percent.png}
    \caption{The estimated factor loads $\factorloads$ (left), factor scores $\factorscores$ (middle),
    		 and specific variances $\specificvariance$ (right) compared to the 
		 	 true data generating values
		 	 for Experiment~2 (Section~\ref{sec:toy-model-missing-data}). Here 40\% of the data are missing at random.
			 The agreement remains excellent,
			 despite the large fraction of missing data entries.}
    \label{fig:exp1-missing-data}
\end{figure*}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{experiments/exp3-\ExperimentHash-latent-factors-and-contributions-with-targets.pdf}
	\caption{Latent factors inferred from 1,072 stars in \Galah\
			 \citep[][thick lines]{Buder:2018} with 17 abundance measurements. Left panels show the absolute entries for each
			 factor load, where the thin lines indicate the target latent factors (see Section~\ref{sec:exp4}). On the right we show the absolute fractional contributions
			 to each element, ordered by the loads that contribute most.}
    \label{fig:exp3-factor-loads}
\end{figure*}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{experiments/exp3-\ExperimentHash-latent-space.pdf}
	\caption{The factor scores $\factorscores$ estimated in Experiment~3 (Section~\ref{sec:exp4} using $N =$ 1,072 stars in the \Galah\ data \citep{Buder:2018} that have 18 abundance measurements. Here each star is coloured by its inferred $K$th component.}
    \label{fig:exp3-latent-space}
\end{figure*}

\begin{figure*}
	\includegraphics[width=\textwidth]{experiments/exp3-\ExperimentHash-data-space.pdf}
	\caption{Detailed chemical abundances from the $N =$ 1,072 stars in \Galah\ \citep{Buder:2018} that have 18 chemical abundances (Section~\ref{sec:exp4}). Each star is coloured by its $K$th inferred component from the lower-dimensional latent space, with the same colouring as per Figure~\ref{fig:exp3-latent-space}.}
    \label{fig:exp3-data-space}
\end{figure*}

@@

\begin{figure*}[t!]
	\includegraphics[width=0.95\textwidth]{experiments/exp10-comparison.pdf}
	\caption{Latent factors found in Section~\ref{sec:exp10} using different subsets
	         of \Galah\ data \citep{Buder:2018}. The thin solid line shows the result from
	         Section~\ref{sec:exp4} with $N=1,072$ stars with 18 abundances and no
	         missing data. Increasing thickness indicates larger samples, up to the solid thick line with $N \sim$ 100,000 stars, where twelve latent factors are preferred.}
    \label{fig:exp10-comparison}
\end{figure*}

@@

This in part constrains our ability to identify new nucleosynthetic processes. For example,
let us consider a hypothetical situation where we would only expect there to be four 
nucleosynthetic processes that predominately contribute to the observed \Galah\ abundances,
but in practice we found that the data are best explained with
five latent factors. We construct a rotation matrix where the first four latent factors
describe the nucleosynthetic processes we expect to find. What of the fifth latent
factor? We can constrain the possible values of the fifth latent factor conditioned on
the requirement that all factors remain mutually orthogonal, but one can imagine that
some (or perhaps many) elements have entries where the fifth latent factor can have
near-zero or zero entries. Even if the mean nucleosynthetic yields are mutually
orthogonal, there are scenarios that one can imagine where there is a limited amount
we can say with confidence about that new nucleosynthetic process.

@@

In the subset of \Galah\ data with complete abundances we find that three components are
preferred. These components can be described as those with (1) low- and (2) high-[$\alpha$/Fe] abundance ratios, and another (3) primarily differing in K, Ba, Zn, and V abundances at a given [Fe/H] and [$\alpha$/Fe] abundance ratio. When we include $\sim$100,000 stars with up to 18 abundances, and assume the incomplete abundances are missing at random, we find that 16 components in latent space are preferred to explain the data. By construction these  components
are structured in their chemical abundances because of the projection from the latent
space, and by extension of each component having similar chemistry, each component
occupies realistic locations in a Hertzsprung-Russell diagram. When we
project these component associations to the data space we find that none of the inferred
components are structured or coherent in their positions or motions. However, in this sample of stars there are no gravitationally bound clusters where a reasonable (e.g. $\sim$30) number of stars have been observed. Clearly, more data would help to resolve a higher number of components.

>>>>

\newcommand{\ExpThreeNumAbundances}{17}

@@

\author[0000-0002-0920-809X]{Sanjib Sharma}						 
\affiliation{Sydney Institute for Astronomy, School of Physics,
			 A28, The University of Sydney, NSW 2006, Australia}

@@

\begin{abstract}
Chemical tagging seeks to identify unique star formation sites from
present-day stellar abundances.
Previous techniques have treated each abundance dimension as being
statistically independent, despite theoretical expectations that
many elements can be produced by more than one nucleosynthetic process.
In this work we introduce a data-driven model of nucleosynthesis 
where a set of latent factors (e.g., nucleosynthetic yields) contribute
to all stars with different scores, and clustering (e.g., chemical tagging) 
is modelled by a mixture of multivariate Gaussians
in a lower-dimensional latent space.
We use an exact method to simultaneously estimate the factor scores for
each star, the partial assignment of each star
to each cluster, and the latent factors common to all stars, even in the
presence of missing data entries.
We use an information-theoretic Bayesian principle to estimate the number of
latent factors and clusters.
Using the second \Galah\ data release we find that five latent factors are
preferred to explain $N =$ 1,072 stars with \ExpThreeNumAbundances\ chemical abundances. 
We identify the rapid- and slow-neutron capture 
processes, as well as latent factors consistent with Fe-peak and 
$\alpha$-element production, and another where only K and Zn dominate.
When we consider $N \sim$ 100,000 stars with missing abundances we find 
another seven factors, as well as 16 components in latent space.
Despite these components showing separation in chemistry that is explained
through different yield contributions, none show significant structure in 
their positions or motions. We argue that more data, and joint priors 
on cluster membership that are constrained by dynamical models, are necessary
to realise chemical tagging at a galactic-scale.
We release accompanying software that scales well with the available data,
allowing for model parameters to be optimised in seconds given a fixed number
of latent factors, components, and $\sim10^7$ abundance measurements.
\end{abstract}

@@

Gravitationally bound star clusters have been useful laboratories for
testing the limits and utility of chemical tagging. Although biases arise when
only considering star clusters that are still gravitationally bound, the chemical
homogeneity of open clusters provides an empirical measure of how similar stars
would need to be before they could be tagged as belonging to the same
star formation site \citep{Bland-Hawthorn:2010b,Bland-Hawthorn:2010a,Mitschang:2014}. However, there are analysis
issues in understanding how precisely those chemical abundances can be measured
\citep{Bovy:2016}, and how chemically similar stars can be that did not form 
together \citep[doppleg\"angers;][]{Ness:2018}.
If open clusters were truely chemically homogeneous then under idealistic 
assumptions our ability to chemically tag the Milky Way would depend primarily
on the precision with which we can measure those chemical abundances in stars. 
Data-driven approaches to modelling stellar spectra are
improving upon this precision \citep{Ness:2015,Ness:2018a,Ness:2018b,
Casey:2016,Casey:2017,Ho:2017b,Ho:2017a,Leung;Bovy:2018,Ting:2019}, but more work is
needed: astronomers have not yet developed unbiased estimators of chemical
abundances that saturate the Cram\'er-Rao bound \citep{Cramer:1946,Rao:1945}.

@@

Given these data and the most favourable assumptions in chemical tagging 
-- that star clusters are truely chemically homogenous, that we can measure 
those abundances with infinite precision, and that those abundances are 
differentiable between star clusters -- then chemical
tagging becomes a clustering problem. All clustering techniques applied to 
chemical tagging thus far have assumed that the data dimensions are independent. That is to say
that adding a dimension of say [Ni/H] provides independent information
that could not have been predicted from other elemental abundances.
Theory and observations agree that this cannot be true.
Nucleosynthetic processes produce multiple elements in varying
quantities, and the effective dimensionality of stellar abundance datasets has been shown
to be lower than the actual number of abundance dimensions \citep{Ting:2012,Price-Jones:2018,Milosavljevic:2018}.
Any clustering approach that treats each new elemental abundance as an 
independent axis of information will therefore conclude with biased inferences
about the star formation history of our Galaxy.

@@

Latent factor analysis is a common statistical approach for describing correlated 
observations with a lower number of latent variables \citep[e.g.,][]{Thompson:2004}.
Related techniques include principal component analysis \citep{Hotelling:1933} and its
variants \citep{Tipping;Bishop:1999}, singular value decomposition \citep{Golub:1970}, and other
matrix factorization methods. While factor analysis on its own is a useful
dimensionality reduction tool to identify latent factors that contribute to
the chemical abundances of stars \citep[e.g.,][]{Ting:2012,Price-Jones:2018,Milosavljevic:2018}, factor
analysis cannot describe clustering in the data (or latent) space. As a result,
some works have performed clustering and then required different latent factors for each (totally assigned) component \citep[e.g.,][]{EdwardsDowe1998}.
Similarly, clustering techniques applied to chemical abundances to date 
\citep[e.g.,][]{Hogg:2016} do not account for the lower effective dimensionality in
elemental abundances.

@@

Here we introduce a toy model where we use generated data to verify that
we recover the true model parameters given some data, and to
ensure that the expectation-maximization method is yielding consistent results.
We generated a data set with ${\NumData = 1}$00,000 data points, each with
$\NumDimensions = 15$ dimensions. We adopted a latent dimensional space of 
$\NumLatentFactors = 5$ factor loads such that the vector $\factorloads$ has
shape $\NumDimensions \times \NumLatentFactors$,
with $\NumComponents = 10$ clusters in the latent space. 
We generated the random factor loads in the same way that we initialise the optimisation (Section~\ref{sec:initialisation}). The relative weights $\vec\weight$
are drawn from a multinomial distribution and the means of the clusters
in factor scores $\scoremeans$ are drawn from a standard normal
distribution. The off-diagonal entries in the covariance matrices in factor scores $\scorecovs$ are drawn from a gamma distribution $\scorecovs_{\numcomponents,i,i} \sim \vec\Gamma\left(1\right)$. The variance in 
each dimension $\specificvariance$ are also drawn $\specificvariance \sim \vec\Gamma\left(1\right)$.
The $\numdata$th data point (which belongs to the $\numcomponents$th cluster) is then
generated by drawing $\factorscores_{\numdata} \sim \mathcal{N}(\scoremeans_\numcomponents,\scorecovs_\numcomponents)$, projecting by the factor loads $\factorloads$, and adding variance $\specificvariance$.

@@

We treat the generated data set as if the number of latent factors
and components are not known. Starting with $\NumLatentFactors = 1$
and $\NumComponents = 1$, we trialled each permutation of $\NumLatentFactors$ and $\NumComponents$
until $\NumLatentFactors_\textrm{max} = 10$
and   $\NumComponents_\textrm{max} = 20$ (e.g., twice the true values of $\NumLatentFactors_\textrm{true}$ and $\NumComponents_\textrm{true}$).

We recorded the \emph{negative} log likelihood, the BIC, and the message length\footnote{Omitting constant terms such that negative message lengths are allowed.}  for each permutation of $\NumLatentFactors$ and $\NumComponents$.
These metrics are shown in Figure~\ref{fig:experiment-1-gridsearch}.
Unsurprisingly the negative log likelihood decreases with increasing numbers of latent
factors $\NumLatentFactors$ and increasing numbers of components $\NumComponents$.
The lowest BIC value and message length is found at $\NumLatentFactors = 5$
and $\NumComponents = 10$, identical to the true values.
It is clear from Figure~\ref{fig:experiment-1-gridsearch} that a combination of latent factors
and clustering in the latent space provides a better description of the (generated) data than a Gaussian mixture model without latent factors.
Adding components to the model does improves the log likelihood, even with a single latent factor,
but the addition of just \emph{one latent factor} improves the log likelihood more so than adding
\emph{twenty components}. Not much more can be said for this example because the true data generating process 
is known, but this toy model does illustrate how 
clustering in high dimensional data can be better described by latent factors with 
clustering in the lower dimensional latent space.

@@

We took the model with the preferred number of latent factors and components found
from a grid search ($\NumComponents = 10$, $\NumLatentFactors = 5$; which are also
the true values) and applied an orthogonal rotation to the latent space to be as
close as possible to the true values. The rotation matrix $\mathbf{R}$ was found
by solving for $\NumLatentFactors$ unknown angle parameters, each of which is used
to construct a Givens rotation matrix \citep{Givens:1958}, and then we take the product of those Givens
matrices to produce a valid rotation matrix $\vec{R}$. This process reduces to Euler angle rotation in three or fewer dimensions.
This process rotates the latent space
($\factorloads$, $\scoremeans$, $\scorecovs$), but has no effect on the model's 
predictive power: the evaluated log likelihood or the Kullback-Leibler divergence \citep{Kullback:1951} under the
rotated model is indistinguishable from the unrotated model.
In Figure~\ref{fig:exp1-compare} we show the estimated factor loads $\factorloads$,
factor scores $\factorscores$, and specific variances $\specificvariance$ compared
to the true values. The agreement is excellent in all model parameters.

@@

\begin{figure*}[t!]
	\includegraphics[width=\textwidth]{experiments/exp-missing-data-not-at-random-inflated-40percent.png}
    \caption{The estimated factor loads $\factorloads$ (left), factor scores $\factorscores$ (middle),
    		 and specific variances $\specificvariance$ (right) compared to the 
		 	 true data generating values
		 	 for Experiment~2 (Section~\ref{sec:toy-model-missing-data}). Here 40\% of the data are missing at random.
			 The agreement remains excellent,
			 despite the large fraction of missing data entries. Note that the scales on the top panels are 2-10 times larger than those in Figure~\ref{fig:exp1-compare}.}
    \label{fig:exp1-missing-data}
\end{figure*}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{experiments/exp3-\ExperimentHash-latent-factors-and-contributions-with-targets.pdf}
	\caption{Latent factors inferred from 1,072 stars in \Galah\
			 \citep[][thick lines]{Buder:2018} with \ExpThreeNumAbundances\ abundance measurements. Left panels show the absolute entries for each
			 factor load, where the thin lines indicate the target latent factors (see Section~\ref{sec:exp4}). On the right we show the absolute fractional contributions
			 to each element, ordered by the loads that contribute most.}
    \label{fig:exp3-factor-loads}
\end{figure*}

@@

\begin{figure*}
	\includegraphics[width=\textwidth]{experiments/exp3-\ExperimentHash-latent-space.pdf}
	\caption{The factor scores $\factorscores$ estimated in Experiment~3 (Section~\ref{sec:exp4} using $N =$ 1,072 stars in the \Galah\ data \citep{Buder:2018} that have \ExpThreeNumAbundances\ abundance measurements. Here each star is coloured by its inferred $K$th component.}
    \label{fig:exp3-latent-space}
\end{figure*}

\begin{figure*}
	\includegraphics[width=\textwidth]{experiments/exp3-\ExperimentHash-data-space.pdf}
	\caption{Detailed chemical abundances from the $N =$ 1,072 stars in \Galah\ \citep{Buder:2018} that have \ExpThreeNumAbundances\ chemical abundances (Section~\ref{sec:exp4}). Each star is coloured by its $K$th inferred component from the lower-dimensional latent space, with the same colouring as per Figure~\ref{fig:exp3-latent-space}.}
    \label{fig:exp3-data-space}
\end{figure*}

@@

\begin{figure*}[t!]
	\includegraphics[width=0.95\textwidth]{experiments/exp10-comparison.pdf}
	\caption{Latent factors found in Section~\ref{sec:exp10} using different subsets
	         of \Galah\ data \citep{Buder:2018}. The thin solid line shows the result from
	         Section~\ref{sec:exp4} with $N=1,072$ stars with \ExpThreeNumAbundances\ abundances and no
	         missing data. Increasing thickness indicates larger samples, up to the solid thick line with $N \sim$ 100,000 stars, where twelve latent factors are preferred.}
    \label{fig:exp10-comparison}
\end{figure*}

@@

This in part constrains our ability to identify new nucleosynthetic processes. For example,
let us consider a hypothetical situation where we would only expect there to be four 
nucleosynthetic processes that predominately contribute to the observed \Galah\ abundances,
but in practice we found that the data are best explained with
five latent factors. We construct a rotation matrix where the first four latent factors
describe the nucleosynthetic processes we expect to find. What of the fifth latent
factor? We can constrain the possible values of the fifth latent factor conditioned on
the requirement that all factors remain mutually orthogonal, but one can imagine that
some (or perhaps many) elements have entries where the fifth latent factor can have
near-zero or zero entries. Even if the mean nucleosynthetic yields are mutually
orthogonal, there are scenarios that one can imagine where there is a limited amount
we can say with confidence about that new nucleosynthetic process \citep[see also][]{Milosavljevic:2018}.

There are similar limitations that arise due to our assumption about the clustering in latent space.
There is no justified reason why the factor scores should be well-described by multivariate normal
distributions. If the \emph{true} underlying scores were not distributed as multivariate normals
then one can imagine similar outcomes when directly fitting data with a mixture of gaussian distributions:
additional components are required to describe complex (non-gaussian) shapes in data space. 
This situation of model mismatch is more extreme when fitting only data rather than the model
described here because some of the data complexity will be described by the orthogonal latent
factors. However, qualitatively the picture is the same: when the \emph{true} underlying distribution
in factor scores are not described by multivariate normals, additional components will likely be
introduced in order to describe non-gaussian features.

@@

In the subset of \Galah\ data with complete abundances we find that three components are
preferred. These components can be described as those with (1) low- and (2) high-[$\alpha$/Fe] abundance ratios, and another (3) primarily differing in K, Ba, Zn, and V abundances at a given [Fe/H] and [$\alpha$/Fe] abundance ratio. When we include $\sim$100,000 stars with up to \ExpThreeNumAbundances\ abundances, and assume the incomplete abundances are missing at random, we find that 16 components in latent space are preferred to explain the data. By construction these  components
are structured in their chemical abundances because of the projection from the latent
space, and by extension of each component having similar chemistry, each component
occupies realistic locations in a Hertzsprung-Russell diagram. When we
project these component associations to the data space we find that none of the inferred
components are structured or coherent in their positions or motions. However, in this sample of stars there are no gravitationally bound clusters where a reasonable (e.g. $\sim$30) number of stars have been observed. Clearly, more data would help to resolve a higher number of components.
